---
title: "Lab08_ReinKatherine"
output: html_document
date: "2024-03-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Data import and processing

```{r}
data <- read.csv("World_Development_Indicators.csv", na.strings = "#N/A")
CO2 = data[,"EN.ATM.CO2E.PC"]
```

## Step 2: Hypotheses generation

1. GDP per capita (current US$)

This one will most likely be positively correlated, as we discussed in class. More money means you are able to spend more on things that emit heavily.

2. Pump price for gasoline (US$ per liter)

This one will most likely be negatively correlated. The higher price of gas would mean people wouldn't want to use a lot of gas meaning CO2 emissions will be lower.

3. Terrestrial and marine protected areas (% of total territorial area)

This one will most likely have a negative correlation. The more area that is protected means that factories or other big emitters can't be on that land.

4. Population density (people per sq. km of land area)

This one will most likely be positively correlated. The more people the more carbon they can emit.

5. Average precipitation in depth (mm per year)

This one will most likely be positively correlated. Carbon emissions can cause a higher amounts of rainfall.

6. Arable land (hectares per person)

This one will most likely have a negative correlation. The more area that is used for farming means that factories or other big emitters can't be on that land.

## Step 3: Exploratory data analysis

```{r}

# Create vectors of actual column names and what we want the variable names to be
xlabel_names = c("GDP","Gas Price","Protected Areas","Population Density",
                   "Precipitation", "Arable Land")
column_names = c("NY.GDP.PCAP.CD", "EP.PMP.SGAS.CD", "ER.PTD.TOTL.ZS", "EN.POP.DNST",
                  "AG.LND.PRCP.MM", "AG.LND.ARBL.HA.PC")

# Pull Out data from dataframe for each variable
GDP = data[,"NY.GDP.PCAP.CD"]
GasPrice = data[,"EP.PMP.SGAS.CD"]
ProtectedAreas = data[,"ER.PTD.TOTL.ZS"]
PopDens = data[,"EN.POP.DNST"]
Precipitation = data[,"AG.LND.PRCP.MM"]
ArableLand = data[,"AG.LND.ARBL.HA.PC"]


# Graph each variable
for (i in 1:6){
  xvariable = column_names[i]
  x_label = xlabel_names[i]
  
  plot(data[,xvariable], CO2, ylab = "Carbon Emissions", xlab = x_label, main = 
         paste(x_label, "vs Carbon Emissions"))
  
  # Fit a linear regression model
  lm_model <- lm(CO2 ~ data[, xvariable])
  
  # Add the line of best fit to the plot
  abline(lm_model, col = "red")
}
```
Precipitation seems to be exponentially related with carbon.

```{r}
# Create table of correlation values
print("***GDP***")
cor(GDP, CO2, use = "complete.obs")
fit_GDP = lm(formula = EN.ATM.CO2E.PC ~ NY.GDP.PCAP.CD, data = data)
summary(fit_GDP)

print("***Gas Price***")
cor(GasPrice, CO2, use = "complete.obs")
fit_GasPrice = lm(formula = EN.ATM.CO2E.PC ~ EP.PMP.SGAS.CD, data = data)
summary(fit_GasPrice)

print("***Protected Areas***")
cor(ProtectedAreas, CO2, use = "complete.obs")
fit_ProtectedAreas = lm(formula = EN.ATM.CO2E.PC ~ ER.PTD.TOTL.ZS, data = data)
summary(fit_ProtectedAreas)

print("***Population Density***")
cor(PopDens, CO2, use = "complete.obs")
fit_PopDens = lm(formula = EN.ATM.CO2E.PC ~ EN.POP.DNST, data = data)
summary(fit_PopDens)

print("***Precipiation***")
cor(Precipitation, CO2, use = "complete.obs")
fit_Precipitation = lm(formula = EN.ATM.CO2E.PC ~ AG.LND.PRCP.MM, data = data)
summary(fit_Precipitation)

print("***Arable Land***")
cor(ArableLand, CO2, use = "complete.obs")
fit_ArableLand = lm(formula = EN.ATM.CO2E.PC ~ AG.LND.ARBL.HA.PC, data = data)
summary(fit_ArableLand)

```

GDP: Yes, I said that they would be positively correlated and they were.
Gas Price: Yes, I said that they would be negatively correlated and they were.
Protected Areas: No, I said negative and there's a very small positive correlation.
Pop Dens: Yes, I said positive and there's a very small positive correlation.
Precipitation: No, I said positive 
Arable Land:



