---
title: "Lab04_ReinKatherine"
output: html_document
date: "2024-02-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A1/A3 - Sustainable Yield Curve

```{r}
# Calculate the growth rate for a vector of N's
log_growth_rate <- function(r,N,K){
  dNdt = r*N*(1-(N/K))
  return(dNdt)
}

# Set Initial Conditions
r = 0.6 # units: fish per year
K = 3200 # units: tonnes
N = 0:K

# Calculate dNdt and graph
dNdt = log_growth_rate(r,N,K)
plot(N,dNdt, type = "l", lwd = 2, ylim = range(dNdt), xlim = range(N), col = "black", main = "Sustainable Yield Curve", xlab = "Population Size (tonnes)", ylab = "Growth Rate (tonnes of fish/year)")

# MSY Point Calculations
equil_stock_size = K/2
msy = (r*K)/4

# Add MSY Point
points(equil_stock_size, msy, pch = 16, col = "red", lwd = 5)

legend("topleft", legend = c("SYC","MSY"), col = c("black","red"), lty = c(1, NA), pch = c(NA, 16), lwd = c(2,5))

```

## A2 - Maximum Sustainable Yield

```{r}
# Set Initial Conditions
r = 0.6 # units: fish per year
K = 3200 # units: tonnes

# Calculate MSY and equilibrium stock size
msy = (r*K)/4
msy
equil_stock_size = K/2
equil_stock_size

```

MSY: 480
Equilibrium stock size: 1600

## B1 - 18% Graph

```{r}
# Set Initial Conditions
r = 0.6 # units: fish per year
K = 3200 # units: tonnes
N = 1:K
N_18 = N

# Calculate dNdt and 18%
dNdt = log_growth_rate(r,N,K)
N_18 = log_growth_rate(r,N*0.18,K)

# Graph
plot(N,dNdt, type = "l", lwd = 2, ylim = range(dNdt), xlim = range(N), col = "black", main = "Sustainable Yield Curve", xlab = "Population Size (tonnes)", ylab = "Growth Rate (tonnes of fish/year)")
lines(N, N_18, lwd = 2, lty = 2, col = "red")
legend("topright", legend = c("SYC","18% SYC"), col = c("black","red"), lty = c(1, 2), lwd = c(2,2))
```

## B2 - Estimate Harvest and Stock Limits

Harvest: 200 tonnes/year
Stock: 2750 tonnes

## B3 - Total Harvest

```{r}
Harvest = 0.18*K
Harvest
```

The first year harvest is 576 which is greater than the MSY. This is possible because the sustainable yield curve simply shows us what is a sustainable harvest not every possible harvest.

## C1 - Liscensed Vessels Graph

```{r}
license_limits <- function(alpha,E,N){
  for(i in 1:length(E)){
    H = alpha*E[i]*N
    lines(N, H, col = i+1, lwd = 2) # color will change with i
  } 
}

# Set Initial Conditions
r = 0.6 # units: fish per year
K = 3200 # units: tonnes
N = 0:K
dNdt = log_growth_rate(r,N,K)
alpha = 0.001
E = c(50,100,200,500) # number of licensed vessels

# Initial Graph
plot(N,dNdt, type = "l", lwd = 2, ylim = range(dNdt), xlim = range(N), col = "black", main = "Sustainable Yield Curve", xlab = "Population Size (tonnes)", ylab = "Growth Rate (tonnes of fish/year)")

# Add License Limits Lines
license_limits(alpha,E,N)
legend("bottomright", title = "Licensed Vessels", legend = E,
         col = 2:(length(E) + 1), lwd = 2)

```

## C2 - Per Vessel Harvest

Tonnes of Fish for...
100 Vessels: 2600 and 260 fish per vessel
500 Vessels: 600 and 1.2 fish per vessel

## C3 - Effort Yield Curve

```{r}
# Initial Conditions
r = 0.6 # units: fish per year
K = 3200 # units: tonnes
N = 0:K
alpha = 0.001

# Define E and H and then graph
Effort_Yield_Curve <- function(r,K,N,alpha){
  effort = (r/alpha)*(1 - (N/K))
  H = alpha*effort*N
  
  # Plot the effort vs harvest vectors
  plot(effort, H, type = "l", lwd = 2, ylim = range(H), xlim = range(effort), 
       col = "black", main = "Effort Yield Curve", xlab = "Effort", ylab 
       ="Harvest")
}

# Call Curve function
Effort_Yield_Curve(r,K,N,alpha)

```

## C4 - Effort Results in MSY

```{r}
# Initialize variables
r = 0.6 # units: fish per year
K = 3200 # units: tonnes
N = 0:K
alpha = 0.001

# Create new vectors
effort = (r/alpha)*(1 - (N/K))
H = alpha*effort*N

# Calculate what effort results in MSY 
optimal_EYC = effort[which.max(H)]
optimal_EYC
```

Effort: 300

## D1 - Revenue, Costs, Profit Graph

```{r}
Revenue_Cost_Graph <- function(r,K,alpha,c,market_price,title){
  # Initial Variables
  N = 0:K
  effort = (r/alpha)*(1 - (N/K))
  H = alpha*effort*N

  # Calculating new vectors
  Total_Cost = effort*c
  Total_Revenue = market_price*H
  Profit = Total_Revenue - Total_Cost

  # Graph
  plot(effort, Total_Revenue, type = "l", lwd = 2, col = "red", ylim = 
         range(Total_Revenue), xlim = c(0,600), main = title, xlab = "Effort",
       ylab = "Revenue ($)")
  lines(effort, Total_Cost, type = "l", lwd = 2, col = "purple")
  lines(effort, Profit, type = "l", lwd = 2, col = "blue")
  legend("bottomright", legend = c("Revenue","Cost","Profit"), col = 
           c("red","purple","blue"), lwd = c(2,2,2))
}

# Initial Conditions
r = 0.6 # units: fish per year
K = 3200 # units: tonnes
alpha = 0.001
c = 10000 # cost per fishing vessel
market_price = 10 * 1000 # cost per kg * tonnes conversion
title = "Revenues and Costs"

# Call function
Revenue_Cost_Graph(r,K,alpha,c,market_price,title)
```

## D2 - Maximize Economic Yield (Effort, Profit, Per Vessel Profit)

```{r}
optimal_effort_values <- function(r,K,alpha,market_price){
  # Calculate new variables
  N = 0:K
  effort = (r/alpha)*(1 - (N/K))
  H = alpha*effort*N

  # Calculating new vectors
  Total_Cost = effort*c
  Total_Revenue = market_price*H
  Profit = Total_Revenue - Total_Cost

  # Calculate Optimal Effort Variables
  optimal_effort_index = which.max(Profit)
  optimal_effort = effort[optimal_effort_index]
  profit_at_OE = Profit[optimal_effort_index]
  per_vessel = profit_at_OE / optimal_effort

  # Print out values
  print(optimal_effort)
  print(profit_at_OE)
  print(per_vessel)
}

# Initial Conditions
r = 0.6 # units: fish per year
K = 3200 # units: tonnes
alpha = 0.001
c = 10000 # cost per fishing vessel
market_price = 10 * 1000 # cost per kg * tonnes conversion

# Print out Optimal Effort Values
optimal_effort_values(r,K,alpha,market_price)

```


Optimal Effort: 206.25
Profit: 2268750
Per Vessel: 11000

## D3 - MEY (Total Harvest)

```{r}
# Initial Conditions
r = 0.6 # units: fish per year
K = 3200 # units: tonnes
alpha = 0.001
c = 10000 # cost per fishing vessel
market_price = 10 * 1000 # cost per kg * tonnes conversion

# Calculate new variables
N = 0:K
effort = (r/alpha)*(1 - (N/K))
H = alpha*effort*N

# Calculating new vectors
Total_Cost = effort*c
Total_Revenue = market_price*H
Profit = Total_Revenue - Total_Cost

# Calculate total harvest
total_harvest_index = which.max(Profit)
total_harvest = H[total_harvest_index]
total_harvest
```

Total Harvest: 433
The total harvest at MEY is double that at MSY.

## D4 - 10% Subsidy

```{r}
# Initial Conditions
r = 0.6 # units: fish per year
K = 3200 # units: tonnes
alpha = 0.001
c = 10000 *.9 # cost per fishing vessel
market_price = 10 * 1000 # cost per kg * tonnes conversion
title = "Revenues and Costs (10% Subsidy)"

# Call function
Revenue_Cost_Graph(r,K,alpha,c,market_price,title)

# Print out Optimal Effort Values
optimal_effort_values(r,K,alpha,market_price)

```
The effort increased to 215. The total profit also increased to 2479688. The profit per vessel increased to 11500.

## D5 - Catchability Increase

```{r}
# Initial Conditions
r = 0.6 # units: fish per year
K = 3200 # units: tonnes
alpha = 0.001 * 1.5 # increase catchability by 50%
c = 10000 # cost per fishing vessel
market_price = 10 * 1000 # cost per kg * tonnes conversion
title = "Revenues and Costs (Catchability Increase)"

# Call function
Revenue_Cost_Graph(r,K,alpha,c,market_price,title)

# Print out Optimal Effort Values
optimal_effort_values(r,K,alpha,market_price)

```

The optimal level of effort decreases to 158. The profit seems to increase to $3008333. The per vessel cost increases to $18995.



